---
title: "K-means Clustering"
author: "Harika"
date: "10/31/2019"
output: html_document
---
Libraries used in this k-means clustering alorithm
```{r setup, include=FALSE}
library(tidyr)
library(tidyselect)
library(factoextra)
library(magrittr)
library(caret)
library(VIM)
library(dplyr)
library(magrittr)
library(imputeTS)
library(stats)
library(flexclust)
```

Given information
The dataset on American College and University Rankings contains information on 1302 American colleges and universities offering an undergraduate program. For each university, there are 17 measurements, including continuous measurements (such as tuition and grad- uation rate) and categorical measurements (such as location by state and whether it is a private or public school).

Question1:1.	Remove all records with missing measurements from the dataset.

```{r}
#read data 
Mydata <- read.csv("Universities.csv")
str(Mydata)

#removing missing measurements from dataset
Mydata_all <- na.omit(Mydata)

Mydata1 <- Mydata_all[-c(1, 2, 3) ]
str(Mydata1)
summary(Mydata1)
summary(Mydata_all)

```
2.	For all the continuous measurements, run K-Means clustering. Make sure to normalize the measurements. How many clusters seem reasonable for describing these data? What was your optimal K?

```{r}
#Data normalization(preprocessing)
set.seed(123)
Mydata1_norm <- preProcess(Mydata1, method=c("center", "scale"))
Mydata1_norm <- predict(Mydata1_norm,Mydata1)
summary(Mydata1_norm)

#To know distance between observations
distance <- get_dist(Mydata1_norm)
fviz_dist(distance)

#K-means clustering algorithm using k=4
k4 <- kmeans(Mydata1_norm, centers = 4, nstart = 25 )
k4

#Finding clustes size,center
k4$cluster
k4$centers
k4$size

#To view results of clutering
fviz_cluster(k4,data= Mydata1_norm)

#k=4 is seems to be reasonable but to find optimal k let us use WSs and silhouette method 
fviz_nbclust(Mydata1_norm,kmeans,method = "wss")
fviz_nbclust(Mydata1_norm,kmeans,method = "silhouette")

#Using ailhouette method it is clearly noticed that k=3 is optimal value
k3 <- kmeans(Mydata1_norm, centers = 3, nstart = 25 )
k3

#finding cluters size and centers
k3$cluster
k3$centers
k3$size

#k=3 is the optimal value 

mydata_new <- data.frame(Mydata1_norm, k3$cluster)
Mydata1 <- data.frame(Mydata1_norm, k3$cluster)
```
3.	Compare the summary statistics for each cluster and describe each cluster in this context (e.g., “Universities with high tuition, low acceptance rate...”).
```{r}

set.seed(123)
k3_dist = kcca(Mydata1_norm, k=3, kccaFamily("kmedians"))
k3_dist
cluster_dist <- predict(k3_dist)
set.seed(123)
cluster<- data.frame(cluster_dist)
Mydata_all <- cbind(Mydata_all, cluster)
Mydata_all <-Mydata_all[-c(22)]
head(Mydata_all)

#combining room,board,add.fees,estimate bookcost,estimate personal cost together
Mydata1$extra_fees <- Mydata1$room + Mydata1$board + Mydata1$add..fees + Mydata1$estim..book.costs + Mydata1$estim..personal..

#Finding summary statistics without private and public column

set.seed(123)
Summary_stats <- Mydata_all %>% group_by(cluster_dist) %>%
 summarise( AcceptanceRate = sum(X..appl..accepted)/ sum(X..appli..rec.d), 
                 new_mean = mean(X..new.stud..enrolled), top10_mean = mean(X..new.stud..from.top.10.), top25_mean = mean(X..new.stud..from.top.25.), out_state_mean = mean(out.of.state.tuition), in_state_mean = mean(in.state.tuition), PHD_mean = mean(X..fac..w.PHD), stufac_mean = mean(stud..fac..ratio), graduation_mean = mean(Graduation.rate))
head(Summary_stats)

```
4.	Use the categorical measurements that were not used in the analysis (State and Private/Public) to characterize the different clusters. Is there any relationship between the clusters and the categorical information?
```{r}
#Summary statistics for State

set.seed(123)
summary_states <-Mydata_all%>% 
             group_by(State) %>% summarise( AcceptanceRate = sum(X..appl..accepted)/ sum(X..appli..rec.d),new_mean=mean(X..new.stud..enrolled),top10_mean=mean(X..new.stud..from.top.10.),top25_mean=mean(X..new.stud..from.top.25.),out_state_mean=mean(out.of.state.tuition), in_state_mean=mean(in.state.tuition), PHD_mean=mean(X..fac..w.PHD), studfac_mean=mean(stud..fac..ratio), gradustio_mean=mean(Graduation.rate), private = sum(Public..1...Private..2. == 2), public = sum(Public..1...Private..2. == 1))
head(summary_states)


#Summary statistics for Private

set.seed(123)
summary_Private <-Mydata_all%>% 
  filter(Public..1...Private..2. == 2) %>% group_by(cluster_dist)  %>% summarise( AcceptanceRate = sum(X..appl..accepted)/ sum(X..appli..rec.d),new_mean=mean(X..new.stud..enrolled),top10_mean=mean(X..new.stud..from.top.10.),top25_mean=mean(X..new.stud..from.top.25.),out_state_mean=mean(out.of.state.tuition), in_state_mean=mean(in.state.tuition), PHD_mean=mean(X..fac..w.PHD), studfac_mean=mean(stud..fac..ratio), gradustio_mean=mean(Graduation.rate))
head(summary_Private)


#Summary statistics for Public
set.seed(123)
summary_Public <-Mydata_all%>% 
  filter(Public..1...Private..2. == 1) %>% group_by(cluster_dist)  %>% summarise( AcceptanceRate = sum(X..appl..accepted)/ sum(X..appli..rec.d),new_mean=mean(X..new.stud..enrolled),top10_mean=mean(X..new.stud..from.top.10.),top25_mean=mean(X..new.stud..from.top.25.),out_state_mean=mean(out.of.state.tuition), in_state_mean=mean(in.state.tuition), PHD_mean=mean(X..fac..w.PHD), studfac_mean=mean(stud..fac..ratio), gradustio_mean=mean(Graduation.rate))
head(summary_Public)

#Cluster 3 has more data points when compared to reamaining clusters.Acceptance rate is low for cluster2.Public universities are more in cluster3.Private university aremore in cluster2 since tuition fees are more
#There is relationship between cluster and categorical variables in the data.

```
Other information like Academic affiliations,Finanacial aid.
6.	Consider Tufts University, which is missing some information. Compute the Euclidean distance of this record from each of the clusters that you found above (using only the measurements that you have). Which cluster is it closest to? Impute the missing values for Tufts by taking the average of the cluster on those measurements. 
```{r}


k3 <- kmeans(Mydata1_norm, centers = 3, nstart = 25 )


Tufts <- filter(Mydata, College.Name == "Tufts University")
#Euclidean distance  from Cluster 1
dist(rbind(Tufts[, -c(1, 2, 3,10)], k3$centers[1,]))
#Euclidean distance  from Cluster 2
dist(rbind(Tufts[, -c(1, 2, 3,10)], k3$centers[2,]))
#Euclidean distance  from Cluster 3
dist(rbind(Tufts[, -c(1, 2, 3,10)], k3$centers[3,]))

#Imputing missing values 
cluster_Tufts <- filter(Mydata_all, cluster_dist == 1)
cluster_Mean <- mean(cluster_Tufts[,c(10)])
Tufts[, c(10)] <- cluster_Mean
cluster_Mean
```






