---
title: "Assignment2.5"
author: "Harika"
date: "10/24/2019"
output: html_document
---
5.	Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on the differences and their reason.

```{r cars}
library(caret)    
library(dplyr)
library(MASS)
library(class)
library(gmodels)
library(dummies)
Data <- read.csv("UniversalBank.csv")
head(Data)
```

```{r}
Data <- Data[,c(-1,-5)]
str(Data)

test <- Data$Education
Test <- factor(test, label =c("E1","E2","E3"))
str(Test)
Data[["Education"]] <- Test
levels(Data$Education)

Dummy<- dummyVars("~Education", data=Data)
x <- predict(Dummy,newdata=Data)

Data <- cbind(Data, x)
Data <- Data[, c(-6)]
```

```{r}
#Data Partition according to the question
set.seed(15)
Test_Index <- createDataPartition(Data$Personal.Loan,p=0.2, list=FALSE) 
Test_Data <- Data[Test_Index,]
TraVal_Data <- Data[-Test_Index,]

Train_Index = createDataPartition(TraVal_Data$Personal.Loan,p=0.625, list=FALSE)
Train_Data = TraVal_Data[Train_Index,]
Valid_data = TraVal_Data[-Train_Index,] 
summary(Train_Data)

#Normalize the data

normalization<-preProcess(Train_Data, method = c("center","scale"))
Train_Norm<-predict(normalization,Train_Data)
Test_Norm<-predict(normalization,Test_Data)
Validation_Norm<-predict(normalization,Valid_data)
All_Data_Norm<-predict(normalization,Data)

Train_Predictors<-Train_Norm[,-7]
Test_predictors <- Test_Norm[,-7]
validation_Predictors<- Validation_Norm[,-7]
All_Data_Predictors<- All_Data_Norm[,-7]

Train_labels <-factor(Train_Data[,7], levels=c(0,1))
Validation_labels  <-factor(Valid_data[,7], levels=c(0,1))
Test_labels <- factor(Test_Data[,7], levels=c(0,1))
All_Data_labels <- factor(Data[,7], levels=c(0,1))
```

```{r}
Predicted_Test_labels <- knn(Train_Predictors, Test_predictors, cl=Train_labels, k=3, prob = TRUE)
str(Predicted_Test_labels)
Predicted_Valid_labels <- knn(Train_Predictors, validation_Predictors, cl=Train_labels, k=3, prob = TRUE)
str(Predicted_Valid_labels)


library(gmodels)
CrossTable(x=Validation_labels, y=Predicted_Valid_labels, prop.chisq = FALSE)

CrossTable(x=Test_labels, y=Predicted_Test_labels, prop.chisq = FALSE)
```

# The accuracy of the model here is reduced when compared to previous data partition because the data given to train is reduced then initially.