---
title: "Naive Bayes Algorithm"
author: "Harika"
date: "10/23/2019"
output: html_document
---
Question:Predicting flight delays can be useful to a variety of organizations. Here,
we look at five predictors. The outcome of interest is whether or not the flight is delayed (delayed here means
arrived more than 15 minutes late). Our data consist of all flights from the Washington, DC area into the New York City area during January 2004. A record is a particular flight. The percentage of delayed flights among these 2201 flights
is 19.5%. The data were obtained from the Bureau of Transportation Statistics (available on the web at www.transtats.bts.gov). The goal is to accurately predict whether or not a new flight (not in this dataset), will be delayed. The outcome variable is whether the flight was delayed, and thus it has two classes (1 = delayed
and 0 = on time).


1.Divide the data into 60% training and 40% validation
```{r}
#Libraries used 
library(caret)
library(e1071)  
library(naivebayes)
library(dplyr)
library(ggplot2)
library(psych)
library(pROC)
library(gmodels)

#Read data file
data <- read.csv("FlightDelays.csv")

#Data preparation
colnames(data)
str(data)
mydata <- data[,c(1,2,4,8,10,13)]

#convert to as factors
mydata$DAY_WEEK <- as.factor(mydata$DAY_WEEK)
mydata$CRS_DEP_TIME <- as.factor(mydata$CRS_DEP_TIME)
str(mydata)

#partition 60% train and 40% validation
set.seed(15)
Index_Train<-createDataPartition(mydata$Flight.Status, p=0.6, list=FALSE)
Train <-mydata[Index_Train,]
Validation  <-mydata[-Index_Train,]

```

2.Run the Naive Bayes model to predict whether the flight is delayed or not. Use only categorical variables for the predictor variables. Note that Week and Time variables need to recode as factors

```{r}
#naive bayes method
model <- naive_bayes(Flight.Status ~ ., data =Train, usekernel = T)
model
```
3.Output both a counts table and a proportion table outlining how many and what proportion of flights were delayed and on-time at each of the three airports.

```{r}
# Finding count and proportion table using piping 

Table_dest <-  Train %>% count(DEST, Flight.Status) %>% mutate(proportion_table = prop.table(n))
Table_dest

#Using table and prop.table functions
counttable <- table(Train$Flight.Status,Train$DEST)
counttable
proportiontable <- prop.table(counttable)
proportiontable

```
4.	Output the confusion matrix and ROC for the validation data

```{r}
#Confusion matrix for training set
Predicted_Test_labels <-predict(model,Train)
confusionMatrix(table(Predicted_Test_labels,Train$Flight.Status))

#Confusion matrix for validation set
Predicted_Test_labels2 <-predict(model,Validation)
confusionMatrix(table(Predicted_Test_labels2,Validation$Flight.Status))
CrossTable(x=Validation$Flight.Status,y=Predicted_Test_labels2, prop.chisq = FALSE) 

model2 <- naiveBayes(Flight.Status ~.,data = Train)
Predicted_Test_labels1 <-predict(model2,Validation, type = "raw")
head(Predicted_Test_labels1)

#ROC
#first column in Predicted_Test_labels1 is delayed status.
roc(Validation$Flight.Status, Predicted_Test_labels1[,1])
plot.roc(Validation$Flight.Status, Predicted_Test_labels1[,1])
```

