---
title: "Project"
author: "Harika"
date: "04/12/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
```{r}
library(factoextra)
library(hrbrthemes)
library(GGally)
library(viridis)
```

Problem Statement:
CRISA has traditionally segmented markets on the basis of purchaser demographics. They would now like to segment the market based on two key sets of variables more directly related to the purchase process and to brand loyalty:
1. Purchase behavior (volume, frequency, susceptibility to discounts, and brand loyalty)
2. Basis of purchase (price, selling proposition)Doing so would allow CRISA to gain information about what demographic attributes are associated with different purchase behaviors and degrees of brand loyalty, and thus deploy promotion budgetsmore effectively. More effective market segmentation would enable CRISA’s clients (in this case, a firm called IMRB)to design more cost-effective promotions targeted at appropriate segments. Thus, multiple promotions could be launched, each targeted at different market segments at different times of the year. This would result in a more cost-effective allocation of the promotion budget to different market segments. It would also enable IMRB to design more effective customer reward systems and thereby increase brand loyalty.

#Reading the data 
```{r}
data <- read.csv("BathSoap.csv")
mydata_norm <- as.data.frame(scale(data))
```

#Brand_Loyal is calculated by the maximum value of all eight brands and comparing with other999 brand.
```{r}
BrandLoyal<-data[,23:30]
Max_Brand<-apply(BrandLoyal, 1, max)
Max_Brand<-as.data.frame(Max_Brand)
BrandComp<-cbind(Max_Brand,data[,31])
MaxLoyal<-as.data.frame(ifelse(BrandComp[,1]>=BrandComp[,2],1,0))
MaxLoyalComp<-cbind(BrandComp,MaxLoyal)
BathSoapComp<-cbind(data,MaxLoyal)
names(BathSoapComp)[47] <- "Brand_Loyal"
```

#1.(a)Purchase Behavior
Purchase behavior variables includes volume, frequency, susceptibility to discounts, and brand loyalty
```{r}
#Variable selection
data1 <- BathSoapComp[,c(12:22,31,47)]
data1 <- as.data.frame(scale(data1))

#k-means
set.seed(123)
data1_K3 <- kmeans(data1, centers =3 , nstart = 200) # k = 3

# Size and Center for the Clusters
data1_K3$size
data1_clusters <- data.frame(data1_K3$centers)
cluster <- matrix(c("1","2","3"),nrow = 3)
data1.visual <- cbind(cluster,data1_clusters)

#Cluster visiualization 
fviz_cluster(data1_K3, data = data1)

#Parallel Plot
ggparcoord(data1.visual,
    columns = 2:14, groupColumn = 1, 
    showPoints = TRUE, 
    title = "Parallel Plot for Purchase Behavior",
    alphaLines = 0.3
    ) 

```
when K=2 the clusters are 
CLuster1:n=223 High loyal customers,favoring no.of brand,brand runs low,Others 999,promotion 6% are low.All are other parameters like Total volume,Trans.brandRuns,No.of trans are high values.
Cluster2:n=377 Low Brand Loyal customers with characteristics opposite to cluster1.

when K=3 the clusters are
Cluster1:n=200 High Loyal customers,favoring low no.of brands,brand runs,avr.price,no promotion,others999.Mid range of ohter promotions.High Volume transaction.
Cluster2:n=237 Low brand loyal customers,favoring towards different promotion types and others 999 brands more.
Cluter3:n=163:It is having customers with high No of brands,brand runs,total volume.Middling all other factors.

#(b)Basis of purchase
Basis of purchase (price, selling proposition)
```{r}

data2 <- BathSoapComp[,c(32:46)]
data2 <- data.frame(scale(data2))
set.seed(123)
data2_k3 <- kmeans(data2, centers =3 , nstart = 500)


# Size and Center for the Clusters
data2_k3$size
data2_clusters <- data.frame(data2_k3$centers)
cluster <- matrix(c("1","2","3"),nrow = 3)
data2.visual <- cbind(cluster,data2_clusters)

#cluster visualization
fviz_cluster(data2_k3, data = data2)

#Parallel Plot
ggparcoord(data2.visual,
    columns = 2:16, groupColumn = 1, 
    showPoints = TRUE, 
    title = "Parallel Plot for Basis of Purchase",
    alphaLines = 0.3
    ) 

```
When k=2
Cluster1:n=78 which has Price Cat:1,2,4 and prop cat 4 to 13 ,15 are low and Price catagory 3,prop cat.14 are high 
Cluster2 :n=522 which shows excatly opposite to cluster1 
In this k=2 analysis it is unclear so about customers trend in most of the parameters.

when k=3
Cluster1:n=376 which has Pr.cat.2 ,4 high.Middling all other values
Cluster2:n= 79 which has characteristics price cat 3 and Prop.Cat14 high.All other are low.
Cluster3:n=145 which has Pr.cat.1,Prop.Cat.6,10,12,13 high volume.



#(c)The variables that describe both purchase behavior and basis of purchase

```{r}

data3 <- BathSoapComp[,c(12:22,32:46,47)]
data3 <- data.frame(scale(data3))

#k-means
set.seed(123)
fviz_nbclust(data3, kmeans, method = "wss")

set.seed(123)
data3_k3 <- kmeans(data3, centers =3 , nstart = 500) 

# Size and Center for the Clusters
data3_k3$size
data3_clusters <- data.frame(data3_k3$centers)
cluster <- matrix(c("1","2","3"),nrow = 3)
data3.visual <- cbind(cluster,data3_clusters)

#cluster visualization
fviz_cluster(data3_k3, data = data3)

#Parallel Plot
ggparcoord(data3.visual,
    columns = 2:13, groupColumn = 1, 
    showPoints = TRUE, 
    title = "Parallel Plot for Purchase Behavior and Basis of PUrchase",
    alphaLines = 0.3
    ) 
ggparcoord(data3.visual,
    columns = 14:28, groupColumn = 1, 
    showPoints = TRUE, 
    title = "Parallel Plot for Purchase Behavior and Basis of PUrchase",
    alphaLines = 0.3
    ) 

```
whenk=3
Cluster3:shows high brand loyal customers and their characteristics
Cluster2:Low brand loyal customers and their characteristics.

# Including Demographics 
```{r}

data4 <- BathSoapComp[,c(2:11)]
data4 <- data.frame(scale(data4))

#k-means

set.seed(123)
data4_k3 <- kmeans(data4, centers =3 , nstart = 500) 

# Size and Center for the Clusters
data4_k3$size
data4_clusters <- data.frame(data4_k3$centers)
cluster <- matrix(c("1","2","3"),nrow = 3)
data4.visual <- cbind(cluster,data4_clusters)

#cluster visualization
fviz_cluster(data4_k3, data = data4)

#Parallel Plot
ggparcoord(data4.visual,
    columns = 2:11, groupColumn = 1, 
    showPoints = TRUE, 
    title = "Parallel Plot for Demographics",
    alphaLines = 0.3
    ) 

```

Each segmentation has it own significance depending upon the companies marketing approaches.Since demographics has been traditional segmentation,Purchase behavior can be helpful for promotion campaigns if company is looking for customers to have trusting relationship that is brand loyalty.









